---
title: "SPARS"
subtitle: "Experiment 1 -- Participants: Descriptive statistics"
author: "Peter Kamerman"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    keep_md: true
    theme: yeti
    hightlight: pygments
    toc: true
    toc_depth: 5
    toc_float: true
    code_folding: show
---

```{r setup, include = FALSE}
# Load packages
library(magrittr)
library(tidyverse) 
library(patchwork) # For multiplot layouts

# knitr setup
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.retina = 2,
                      fig.align = 'center',
                      fig.path = 'figures/1A-participants-descriptive/')
```

----

This script generates descriptive statistics of the SPARS (trial A) cohort, and includes assessments of:

- Age  
- Sex  
- Pain Catastrophizing Scale (PCS)  
- Positive and Negative Affect Schedule (PANAS)  
- Depression, Anxiety, and Stress Scale (DASS-42)

----

# Import and inspect data
```{r import}
# Import
data <- read_rds('./data-cleaned/SPARS_A.rds')

# Inspect
dim(data)
names(data)
head(data)
tail(data)
glimpse(data)
```

----

# Clean data
```{r clean_data}
# Select participant descriptive data
data %<>%
  select(PID, age:pcs_helplessness)

# Filter unique data for each individual
data %<>%
  group_by(PID) %>%
  distinct() %>%
  ungroup()

# Quick inspection for duplicates
data

# Remove ID15 duplicate (row 16, one miscoded value of 13 for dass42_anxiety)
data <- data[-16, ]
```

----

# Age 

```{r age, fig.width = 3.5}
# Tabulate
data %>%
  select(age) %>%
  summarise(Mean = round(mean(age), 1),
            Median = median(age),
            iqr = paste(round(quantile(age, 0.25)), '-', 
                        round(quantile(age, 0.75))),
            range = paste(min(age), '-', max(age))) %>%
  knitr::kable(x = .,
               caption = 'Summary table of age in years',
               col.names = c('Mean', 'Median', 'IQR', 'Range'),
               align = 'rrrr')

# Plot
data %>%
  select(age) %>%
  ggplot(data = .) +
  aes(x = 'Age',
      y = age) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  labs(title = "Summary plot of participants' ages",
       y = 'Age (years)') +
  theme_bw() +
  theme(axis.title.x = element_blank())
```

----

# Sex

```{r sex, fig.width = 3.5}
# Process data
sex <- data %>%
  select(sex) %>%
  mutate(sex = factor(sex),
         sex = fct_recode(sex, 
                          male = '1',
                          female = '2')) %>%
  group_by(sex) %>%
  summarise(count = n(),
            proportion = round(count / nrow(.), 2))

# Tabulate data
knitr::kable(x = sex,
             caption = 'Summary table sex ratio',
             col.names = c('', 'Count', 'Proportion'),
             align = 'lrr')

# Plot
ggplot(data = sex) +
  aes(x = 'sex',
      y = count,
      fill = sex) +
  geom_col(position = position_fill()) +
  scale_fill_grey(start = 0.4) +
  labs(title = "Summary plot of sex ratio",
       y = 'Proportion') +
  theme_bw() +
  theme(axis.title.x = element_blank())
```

----

# Pain Catastrophizing Scale (PCS)

The Pain Catastrophizing Scale (PCS) [^0] is a 13-item scale, with each item rated on a 5-point Likert scale: 0 (not at all) to 4 (all the time). The PCS consists of three subscales being magnification (3 items), rumination (4 items), and helplessness (6 items).

[^0]: Sullivan M, Bishop SR, Pivik J. The pain catastrophizing scale: development and validation. _Psychol. Assess._ **7**: 524, 1995. [Download](http://sullivan-painresearch.mcgill.ca/pdf/abstracts/sullivanapr1995.pdf)

```{r pcs, fig.height = 8}
# Process data 
pcs <- data %>%
  select(starts_with('pcs')) %>%
  mutate(pcs_total = rowSums(.)) 

# Tabulate
pcs %>%
  gather(key = subscale,
         value = score) %>%
  group_by(subscale) %>%
  summarise(Mean = round(mean(score), 1),
            Median = median(score),
            iqr = paste(round(quantile(score, 0.25)), '-', 
                        round(quantile(score, 0.75))),
            range = paste(min(score), '-', max(score))) %>%
  knitr::kable(x = .,
               caption = 'Summary table of PCS total and subscale scores',
               col.names = c('', 'Mean', 'Median', 'IQR', 'Range'),
               align = 'lrrrr')

# Plot
total <- pcs %>%
  select(pcs_total) %>%
  ggplot(data = .) +
  aes(x = 'Total',
      y = pcs_total) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  geom_hline(yintercept = 30,
             linetype = 2) +
  lims(y = c(0, 52)) +
  labs(title = 'Summary plot of PCS total and subscale scores',
       subtitle = 'Total score (normal < 30)',
       y = 'Score [0 - 52]') +
  theme_bw() +
  theme(axis.title.x = element_blank())

rumination <- pcs %>%
  select(pcs_rumination) %>%
  ggplot(data = .) +
  aes(x = 'Rumination',
      y = pcs_rumination) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  lims(y = c(0, 16)) +
  labs(title = '',
       subtitle = 'Rumination score\n(max score = 16)',
       y = 'Score [0 - 16]') +
  theme_bw() +
  theme(axis.title.x = element_blank())

magnification <- pcs %>%
  select(pcs_magnification) %>%
  ggplot(data = .) +
  aes(x = 'Magnification',
      y = pcs_magnification) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  scale_y_continuous(breaks = c(0, 3, 6, 9, 12), limits = c(0, 12)) +
  labs(title = '',
       subtitle = 'Magnification score\n(max score = 12)',
       y = 'Score [0 - 12]') +
  theme_bw() +
  theme(axis.title.x = element_blank())

helplessness <- pcs %>%
  select(pcs_helplessness) %>%
  ggplot(data = .) +
  aes(x = 'Helplessness',
      y = pcs_helplessness) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  lims(y = c(0, 24)) +
  labs(title = '',
       subtitle = 'Helplessness score\n(max score = 24)',
       y = 'Score [0 - 24]') +
  theme_bw() +
  theme(axis.title.x = element_blank())

# Patch plots together
total + {magnification + rumination + helplessness} + plot_layout(nrow = 2)
```

----

# Positive and Negative Affect Schedule (PANAS)

The Positive and Negative Affect Schedule (PANAS) [^1] is a self-report questionnaire that consists of two 10-item scales to measure positive and negative affect. Each item is rated on a Likert scale of 1 (not at all) to 5 (very much).

[^1]: Crawford JR, Henry JD. The positive and negative affect schedule (PANAS): construct validity, measurement properties and normative data in a large non-clinical sample. _Br J Clin Psychol_ **43**: 245-65, 2004. PMID: [15333231](https://www.ncbi.nlm.nih.gov/pubmed/15333231).

```{r panas}
# Process
panas <- data %>%
  select(PID, starts_with('panas')) %>%
  gather(key = affect,
         value = score,
         -PID) %>%
  mutate(total_score = as.integer(score),
         affect = fct_recode(affect, 
                             Negative = 'panas_negative',
                             Positive = 'panas_positive'))

# Tabulate
panas %>%   
  group_by(affect) %>%
  summarise(median = median(score),
            iqr = paste(round(quantile(score, 0.25)), '-', 
                        round(quantile(score, 0.75))),
            range = paste(min(score), '-', max(score))) %>%
  knitr::kable(x = .,
               caption = 'Summary table of PANAS subscale scores',
               col.names = c('Affect', 'Median', 'IQR', 'Range'),
               align = 'lrrr')

# Plot
ggplot(data = panas) +
  aes(x = affect,
      y = score,
      fill = affect) +
  geom_boxplot() +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  scale_fill_grey(start = 0.4) +
  lims(y = c(10, 50)) +
  labs(title = 'Summary plot of PANAS subscale scores',
       x = 'Affect',
       y = 'Score [10 - 50]',
       subtitle = 'Negative scale: lower values are better\nPositive Scale: greater values are better') +
  theme_bw()
```

----

# Depression, Anxiety and Stress Scale (DASS-42)

The Depression Anxiety Stress Scales [^2] is a 42 self-report item scale assessing negative emotional symptoms. Each item is rated on a four-point Likert scale of frequency or severity of the participants' experiences over the last week with the intention of emphasising states over traits. These scores ranged from 0 (did not apply at all) to 4 (applied to them very much/most of the time).

[^2]: Crawford JR, Henry JD. The Depression Anxiety Stress Scales (DASS): normative 
data and latent structure in a large non-clinical sample. _Br J Clin Psychol_ **42**: 111-31, 2003. PMID: [12828802](https://www.ncbi.nlm.nih.gov/pubmed/12828802)

```{r dass}
# Tabulate
data %>%
  select(PID, dass42_depression, dass42_anxiety, dass42_stress) %>%
  rename(depression = dass42_depression,
         anxiety = dass42_anxiety,
         stress = dass42_stress) %>%
  gather(key = dimension,
         value = score,
         -PID) %>%
  group_by(dimension) %>%
  summarise(median = median(score),
            iqr = paste(round(quantile(score, 0.25)), '-', 
                        round(quantile(score, 0.75))),
            range = paste(min(score), '-', max(score))) %>%
  knitr::kable(x = .,
               caption = 'Summary table of DASS-42 dimension scores',
               col.names = c('Dimension', 'Median', 'IQR', 'Range'),
               align = 'lrrr')
  
# Process plots
depression <- data %>%
  select(PID, dass42_depression) %>%
  rename(depression = dass42_depression) %>%
  ggplot(data = .) +
  aes(x = 'Depression',
      y = depression) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  geom_hline(yintercept = 9, 
             linetype = 2) +
  lims(y = c(0, 56)) +
  labs(title = 'Summary plot of DASS-42',
       subtitle = 'Depression (normal < 10)',
       y = 'Score [0 - 56]') +
  theme_bw() +
  theme(axis.title.x = element_blank())

anxiety <- data %>%
  select(PID, dass42_anxiety) %>%
  rename(anxiety = dass42_anxiety) %>%
  ggplot(data = .) +
  aes(x = 'Anxiety',
      y = anxiety) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  geom_hline(yintercept = 7, 
             linetype = 2) +
  lims(y = c(0, 56)) +
  labs(title = '',
       subtitle = 'Anxiety (normal < 8)',
       y = '') +
  theme_bw() +
  theme(axis.title.x = element_blank())

stress <- data %>%
  select(PID, dass42_stress) %>%
  rename(stress = dass42_stress) %>%
  ggplot(data = .) +
  aes(x = 'Stress',
      y = stress) +
  geom_boxplot(outlier.size = -1,
               fill = '#cccccc') +
  geom_point(shape = 21,
             size = 2,
             fill = '#FFFFFF',
             position = position_jitter(height = 0)) +
  geom_hline(yintercept = 14, 
             linetype = 2) +
  lims(y = c(0, 56)) +
  labs(title = '',
       subtitle = 'Stress (normal < 15)',
       y = '') +
  theme_bw() +
  theme(axis.title.x = element_blank())

# Patch them together
depression + anxiety + stress + plot_layout(ncol = 3)
```

----

# Session information
```{r session_info}
sessionInfo()
```
