---
title: "Supplement 1"
subtitle: "Experiment 2 -- Instructions on how to use each rating scale"
author: "Tory Madden"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
      toc: true
      toc_depth: 2
      latex_engine: xelatex
mainfont: Helvetica
fontsize: 12pt
geometry: a4paper, margin=15mm
---

```{r setup, include = FALSE}
# knitr setup
knitr::opts_chunk$set(fig.align = 'center')
```

----

This document displays the information given to participants before each block of Experiment 2. 

----

# Rating scale

## Sensation and Pain Rating Scale (SPARS)

\begin{figure}[h]
\centering
\includegraphics{supplementary-files/spars.png}
\end{figure}

\pagebreak

## Numerical Pain Rating Scale (NRS)

\begin{figure}[h]
\centering
\includegraphics{supplementary-files/nrs.png}
\end{figure}

## Numerical Sensation Rating Scale (SRS)

\begin{figure}[h]
\centering
\includegraphics{supplementary-files/srs.png}
\end{figure}

----

# Session information

```{r session_info}
sessionInfo()
```
